---
title: "R_Assignment_WAB"
output:
  pdf_document: default
  html_document: default
---

R Assignment


--- installs packages from tidyverse
--- reads fang genotypes file into data frame
--- reads snp positions into data frame
--- selects SNP_ID, Chromosome, and Position columns to new data frame
--- filters genotypes to extract maize genotype data to new data frame, str_detect detects presence of match for desired specimen groups, filters those
--- filters genotypes to extract teosinte genotype data to new data frame, str_detect detects presence of match for desired specimen groups, filters those
--- removes header info from maize snps, transposes, defines as data frame, turns row names into columns, renames as SNP_ID to facilitate joining to snps
--- removes header info from teosinte snps, transposes, defines as data frame, turns row names into columns, renames as SNP_ID to facilitate joining to snps
--- merges SNP info with maize genotype data
--- merges SNP info with teosinte genotype data

--- sorts maize SNPs by chromosome, then position ascending, numerically
--- sorts teosinte SNPs by chromosome, then position ascending, numerically
--- sorts maize SNPs by chromosome, then position descending, numerically to separate df 
--- sorts teosinte SNPs by chromosome, then position descending, numerically to separate df 
--- replaces ?/? with -/- in descending maize SNPs file
--- replaces ?/? with -/- in descending teosinte SNPs file


save(maize_1, file="Result_Files/maize_1.RData")


  
  working order 
  - split into files and print to .txt for each maize/teosinte ascending/descending
 - visualize
 - data inspection
  
for visualization, have one file with teosinte/maize snps, including group info? filter as before for maize/teo groups into one file, transpose (-(Sample_ID:JG_OTU)), merge, filter unknown/multiple - can visualize by group - make sure group column named as group if making one large file

removed from workflow - unnecessary filtering? - can use final 40 files to graph

--- removes SNPs with multiple or unknown positions from maize SNPs
--- removes SNPs with multiple or unknown positions from teosinte SNPs

maize_snps <- maize_snps %>% filter(!(Chromosome %in% c("multiple", "unknown") | Position %in% c("multiple", "unknown")))
teosinte_snps <- teosinte_snps %>% filter(!(Chromosome %in% c("multiple", "unknown") | Position %in% c("multiple", "unknown")))
  
```{r}
rm(list = ls())
library(tidyverse)
fang_genos <- read_tsv("Data/fang_et_al_genotypes.txt")
snp_pos <- read_tsv("Data/snp_position.txt")
snps <- select(snp_pos, SNP_ID, Chromosome, Position)
maize_genos <- fang_genos %>% 
    filter(str_detect(Group, "ZMMIL|ZMMLR|ZMMMR"))
teosinte_genos <- fang_genos %>% 
    filter(str_detect(Group, "ZMPBA|ZMPIL|ZMPJA"))
maize_transposed <- rownames_to_column(as.data.frame(t(select(maize_genos, -(Sample_ID:Group)))), var = "SNP_ID")
teosinte_transposed <- rownames_to_column(as.data.frame(t(select(teosinte_genos, -(Sample_ID:Group)))), var = "SNP_ID")
maize_snps <- merge(snps, maize_transposed)
teosinte_snps <- merge(snps, teosinte_transposed)
maize_snps <- arrange(maize_snps, as.numeric(Chromosome), as.numeric(Position))
teosinte_snps <- arrange(teosinte_snps, as.numeric(Chromosome), as.numeric(Position))
maize_snps_desc <- arrange(maize_snps, as.numeric(Chromosome), desc(as.numeric(Position)))
teosinte_snps_desc <- arrange(teosinte_snps, as.numeric(Chromosome), desc(as.numeric(Position)))
maize_snps_desc <- maize_snps_desc %>% mutate(across(everything(), ~ gsub("\\?/\\?", "-/-", .)))
teosinte_snps_desc <- teosinte_snps_desc %>% mutate(across(everything(), ~ gsub("\\?/\\?", "-/-", .)))

```
Separates maize SNP data into data frames for each chromosome, both ascending and descending by position, saves as .RData files in /Result_Files folder
```{r}
maize_1 <- filter(maize_snps, Chromosome == 1)
save(maize_1, file="Result_Files/maize_1.RData")
maize_2 <- filter(maize_snps, Chromosome == 2)
save(maize_2, file="Result_Files/maize_2.RData")
maize_3 <- filter(maize_snps, Chromosome == 3)
save(maize_3, file="Result_Files/maize_3.RData")
maize_4 <- filter(maize_snps, Chromosome == 4)
save(maize_4, file="Result_Files/maize_4.RData")
maize_5 <- filter(maize_snps, Chromosome == 5)
save(maize_5, file="Result_Files/maize_5.RData")
maize_6 <- filter(maize_snps, Chromosome == 6)
save(maize_6, file="Result_Files/maize_6.RData")
maize_7 <- filter(maize_snps, Chromosome == 7)
save(maize_7, file="Result_Files/maize_7.RData")
maize_8 <- filter(maize_snps, Chromosome == 8)
save(maize_8, file="Result_Files/maize_8.RData")
maize_9 <- filter(maize_snps, Chromosome == 9)
save(maize_9, file="Result_Files/maize_9.RData")
maize_10 <- filter(maize_snps, Chromosome == 10)
save(maize_10, file="Result_Files/maize_10.RData")

maize_1_desc <- filter(maize_snps_desc, Chromosome == 1)
save(maize_1_desc, file="Result_Files/maize_1_desc.RData")
maize_2_desc <- filter(maize_snps_desc, Chromosome == 2)
save(maize_2_desc, file="Result_Files/maize_2_desc.RData")
maize_3_desc <- filter(maize_snps_desc, Chromosome == 3)
save(maize_3_desc, file="Result_Files/maize_3_desc.RData")
maize_4_desc <- filter(maize_snps_desc, Chromosome == 4)
save(maize_4_desc, file="Result_Files/maize_4_desc.RData")
maize_5_desc <- filter(maize_snps_desc, Chromosome == 5)
save(maize_5_desc, file="Result_Files/maize_5_desc.RData")
maize_6_desc <- filter(maize_snps_desc, Chromosome == 6)
save(maize_6_desc, file="Result_Files/maize_6_desc.RData")
maize_7_desc <- filter(maize_snps_desc, Chromosome == 7)
save(maize_7_desc, file="Result_Files/maize_7_desc.RData")
maize_8_desc <- filter(maize_snps_desc, Chromosome == 8)
save(maize_8_desc, file="Result_Files/maize_8_desc.RData")
maize_9_desc <- filter(maize_snps_desc, Chromosome == 9)
save(maize_9_desc, file="Result_Files/maize_9_desc.RData")
maize_10_desc <- filter(maize_snps_desc, Chromosome == 10)
save(maize_10_desc, file="Result_Files/maize_10_desc.RData")
```
Separates teosinte SNP data into data frames for each chromosome, both ascending and descending by position, saves as .RData files in /Result_Files folder
```{r}
teosinte_1 <- filter(teosinte_snps, Chromosome == 1)
save(teosinte_1, file="Result_Files/teosinte_1.RData")
teosinte_2 <- filter(teosinte_snps, Chromosome == 2)
save(teosinte_2, file="Result_Files/teosinte_2.RData")
teosinte_3 <- filter(teosinte_snps, Chromosome == 3)
save(teosinte_3, file="Result_Files/teosinte_3.RData")
teosinte_4 <- filter(teosinte_snps, Chromosome == 4)
save(teosinte_4, file="Result_Files/teosinte_4.RData")
teosinte_5 <- filter(teosinte_snps, Chromosome == 5)
save(teosinte_5, file="Result_Files/teosinte_5.RData")
teosinte_6 <- filter(teosinte_snps, Chromosome == 6)
save(teosinte_6, file="Result_Files/teosinte_6.RData")
teosinte_7 <- filter(teosinte_snps, Chromosome == 7)
save(teosinte_7, file="Result_Files/teosinte_7.RData")
teosinte_8 <- filter(teosinte_snps, Chromosome == 8)
save(teosinte_8, file="Result_Files/teosinte_8.RData")
teosinte_9 <- filter(teosinte_snps, Chromosome == 9)
save(teosinte_9, file="Result_Files/teosinte_9.RData")
teosinte_10 <- filter(teosinte_snps, Chromosome == 10)
save(teosinte_10, file="Result_Files/teosinte_10.RData")

teosinte_1_desc <- filter(teosinte_snps_desc, Chromosome == 1)
save(teosinte_1_desc, file="Result_Files/teosinte_1_desc.RData")
teosinte_2_desc <- filter(teosinte_snps_desc, Chromosome == 2)
save(teosinte_2_desc, file="Result_Files/teosinte_2_desc.RData")
teosinte_3_desc <- filter(teosinte_snps_desc, Chromosome == 3)
save(teosinte_3_desc, file="Result_Files/teosinte_3_desc.RData")
teosinte_4_desc <- filter(teosinte_snps_desc, Chromosome == 4)
save(teosinte_4_desc, file="Result_Files/teosinte_4_desc.RData")
teosinte_5_desc <- filter(teosinte_snps_desc, Chromosome == 5)
save(teosinte_5_desc, file="Result_Files/teosinte_5_desc.RData")
teosinte_6_desc <- filter(teosinte_snps_desc, Chromosome == 6)
save(teosinte_6_desc, file="Result_Files/teosinte_6_desc.RData")
teosinte_7_desc <- filter(teosinte_snps_desc, Chromosome == 7)
save(teosinte_7_desc, file="Result_Files/teosinte_7_desc.RData")
teosinte_8_desc <- filter(teosinte_snps_desc, Chromosome == 8)
save(teosinte_8_desc, file="Result_Files/teosinte_8_desc.RData")
teosinte_9_desc <- filter(teosinte_snps_desc, Chromosome == 9)
save(teosinte_9_desc, file="Result_Files/teosinte_9_desc.RData")
teosinte_10_desc <- filter(teosinte_snps_desc, Chromosome == 10)
save(teosinte_10_desc, file="Result_Files/teosinte_10_desc.RData")
```

